---
title: 'WORKSHEET #6'
author: "Ann Erika D. Gabales"
date: "2022-12-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1.How many columns are in mpg dataset? How about the number of rows? Show the codes and its result.
```{r }
library(ggplot2)
library(dplyr)

data(mpg)
nrow(mpg)
ncol(mpg)
```
  **There are 11 columns and 234 rows in mpg data set.**
  
## 2. Which manufacturer has the most models in this data set? Which model has the most variations? Ans:
```{r }
manu_mpg <- mpg$manufacturer
table(manu_mpg)
```
  **Dodge has the most models and variations in this data set.**

### 2.a Group the manufacturers and find the unique models. Copy the codes and result.
```{r }
group_data1 <- mpg

unique_mod <- group_data1 %>% group_by(manufacturer, model) %>%
  distinct() %>% count()
unique_mod

colnames(unique_mod) <- c("Manufacturer", "Model", "Counts")
unique_mod
```

### 2.b Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r }
qplot(model, data = mpg,geom = "bar", fill=manufacturer)
ggplot(mpg, aes(model, manufacturer)) + geom_point()
```

## 3. Same dataset will be used. You are going to show the relationship of the model and the manufacturer.
```{r }
rs_data <- mpg
models <- rs_data %>% group_by(manufacturer, model) %>%
  distinct() %>% count()
models
colnames(models) <- c("Manufacturer", "Model", "Counts")
models
```

### 3.a What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r }
ggplot(mpg, aes(model, manufacturer)) + geom_point()
```
**This code shows a geometric point graph of model and manufacturer of the mpg data set.**

### 3.b For you, is it useful? If not, how could you modify the data to make it more informative?

**This plot is useful since it efficiently creates a data visualization for us to have an easy access.**

## 4. Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result.
```{r }
rs_data2 <- models %>% group_by(Model) %>% count()
colnames(rs_data2) <- c("Model","Counts")
rs_data2
```

### 4.a Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r }
qplot(mpg$model, data = mpg,
      main = "Number of Cars per Model",
      xlab = "Model", ylab = "Number of Cars",
      geom = "bar", fill = mpg$manufacturer) + coord_flip()
```

### 4.b Use only the top 20 observations. Show code and results.
```{r }
data_obsvtn <- subset(mpg[c(1:20),c(1:11)])
manu_obsvtn <- mpg$manufacturer[1:20]
model_obsvtn <- mpg$model[1:20]

qplot(model_obsvtn, data = data_obsvtn, main = "Number of Cars per Model", 
      xlab = "Model",ylab = "Number of Cars", geom = "bar", fill = manu_obsvtn) + coord_flip()
```

## 5. Plot the relationship between cyl - number of cylinders and displ - engine displace-ment using geom_point with aesthetic colour = engine displacement. Title should be “Relationship between No. of Cylinders and Engine Displacement”.

### 5.a Show the codes and its result.
```{r }
ggplot(data = mpg, 
       mapping = aes(x = displ, 
                    y = cyl, 
                    main = "Relationship between No. of Cylinders and Engine Displacement")) + 
  geom_point(mapping=aes(colour = "engine displacement")) 
```

### 5.b How would u describe its relationships?

 **The engine displacement for a car is higher for higher number of cylinders. In other words, as the displ increases, the number of cylinders also increases.**
 
## 6. Get the total number of observations for drv - type of drive train (f = front-wheel drive, r = rear wheel drive, 4 = 4wd) and class - type of class (Example: suv, 2seater, etc.) Plot using the geom_tile() where the number of observations for class be used as a fill for aesthetics.

### 6.a Show the codes and its result for the narrative in #6.
```{r }
mpg %>%
  count(class, drv) %>%
  ggplot(aes(x = class, y = drv)) +
    geom_tile(mapping = aes(fill = n))
```

### 6.b Interpret the result.

**The total number of observations for the drv and class type are displayed in the result. The geom_tile() uses a color scale to show the number of observations with each (x, y) value.**

## 7. Discuss the difference between these codes. Its outputs for each are shown below.
```{r }
#Code 1
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, colour = "blue"))

#Code 2
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), colour = "blue")
```
**The difference between Code 1 and Code 2 is quite easy to tell. We can see that in Code 1, the colour was put inside the geom_point() function and it was resulted with a legend, while in Code 2, the colour was outside the geom_point() function but it has no legend unlike Code 1. But the clear difference between the two is that the points in Code 1 are not blue because the “color =” parameter lies within aes(). This means the function will be looking for a column within the mpg dataset called “blue”, which does not exist.**

## 8. Try to run the command ?mpg. What is the result of this command?
```{r }
?mpg
```

### 8.a Which variables from mpg dataset are categorial?

**The variables that are categorical in the mpg data set are manufacturer, model, trans, drv, fl, and class.**

### 8.b Which are continuous variables?

**The continuous variables in the mpg data set are displ, year, cyl.**

### 8.c. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #5-b. What is its result? Why it produced such output?
```{r }
ggplot(mpg, aes(x = displ, y = hwy, colour = cty)) + geom_point()
```

## 9. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon) using geom_point(). Add a trend line over the existing plot using geom_smooth() with se = FALSE. Default method is “loess”.
```{r }
ggplot(data= mpg, mapping = aes(x= displ, y= hwy)) + geom_point(mapping= aes(color= class)) +
  geom_smooth(se= FALSE)
```

## 10.Using the relationship of displ and hwy, add a trend line over existing plot. Set the se = FALSE to remove the confidence interval and method = lm to check for linear modeling
```{r }
ggplot(data= mpg, mapping= aes(x= displ, y= hwy, color= class)) + geom_point() +
  geom_smooth(se= FALSE)
```